---
always_allow_html: yes
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Project Status: Active](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active) 
<!-- badges: end -->

# DepartureTime

R package to generate departure times for accessibility analysis

## Description

The goal of DepartureTime is to prepare a `data.frame` which contains generated departure times, using user-defined:

+ sampling method
+ date
+ time-window
+ path and file name

Supported sampling methods:

+ **Systematic** sampling method: departure times are selected using a regular interval defined by the frequency 
+ **Simple Random** sampling method: a specified number of departure times (defined by the frequency) is randomly selected from the time window
+ **Hybrid** sampling method: departure times are randomly selected from given time intervals (resulted from applied temporal resolution)
+ **Constrained Random Walk Sampling** sampling method: a first departure time is randomly selected from the subset of the length defined by the frequency and beginning of the time window; then, the next departure time is randomly selected from the subset limited by $Tn+f/2$ and $Tn+f+f/2$

For details please consult [Owen & Murphy (2018)](https://trid.trb.org/view/1497217).

## Installation

You can install `DepartureTime` pacakage from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("stmarcin/DepartureTime")
```

## Function syntax:

```
library(DepartureTime)

DepartureTime <- function(method = "H",
                          dy = format(Sys.Date(), "%Y"),  
                          dm = format(Sys.Date(), "%m"), 
                          dd = format(Sys.Date(), "%d"),
                          tmin = 0, tmax = 24,
                          res = 5,
                          MMDD = TRUE,
                          ptw = FALSE)
```

## Function variables:

+ `method` - sampling method; Options:
    + `R` OR `Random`: Simple random sampling;
    + `S` OR `Systematic`: Systematic sampling;
    + `H` OR `Hybrid`: Hybrid sampling;
    + `W` OR `Walk`: Constrained random walk sampling;
+ `dy`, `dm` and `dd` - date of the analysis (formats: YYYY, MM, DD); **default: system date**;
+ `tmin` and `tmax` - limits of the time window (format: HH); **default: full day** (00:00 - 24:00);
+ `res` - temporal resolution; **default: 5 minutes**
+ `MMDD` - date format of the output (TRUE / FALSE) **default: TRUE**
    + `TRUE`: MM/DD/YYYY;
    + `FALSE`: DD/MM/YYYY;
+ `ptw` - print limits of subsetted time-windows; **default: FALSE**;


## Output

`data.frame` which contains generated departure times (to be used e.g. in ArcGIS Network to generated ODs with time-dependent transport data, e.g. GTFS). File structure:

```{r, echo=FALSE, eval = T, warning=F}
library(knitr)
library(kableExtra)
FileStructure <- data.frame(ColumnName = c("ID", "Date"),
                            Description = c("rowID (integer), starts with 0", "Departure date & hour"))

kable_styling(kable(FileStructure), full_width = F)
```


## Examples

Working example, uses all default variables and hybrid sampling method:

```
DepartureTime() %>% 
  head()
```

```{r, echo = F, warning = F, message = F }
library(dplyr)

DepartureTime::DepartureTime() %>% 
  head() %>% 
  knitr::kable(format='html', escape = FALSE) %>% 
  kableExtra::kable_styling(full_width = F)
```


Example with user-defined parameters:
```{r, eval = F}
DepartureTime(method = "S",    # systematic sampling method
  dm = 5, dd = 15,             # user-defined date: 15th May, 2020 (current year)
  tmin = 7, tmax = 9,          # user-defined time window (07:00 - 09:00)
  res = 20)                    # user-defined temporal resolution (20 minutes)
```

```{r, echo = F, warning = F, message = F }
# library(dplyr)

DepartureTime::DepartureTime(method = "S",  dm = 5, dd = 15,  tmin = 7, tmax = 9, res = 20) %>% 
  knitr::kable(format='html', escape = FALSE) %>% 
  kableExtra::kable_styling(full_width = F)
```

## Direct export to `.dbf`

If you don't want/need to revise departure times, you can easily export them directly, e.g. to `.dbf` file using `{foreign}` package:

```{r, eval = F, warning = F, message = F }
library(DepartureTime)
library(foreign)
library(dplyr)

# generate departure times for 8-10am time window 
# with 30-minute temporal resolution applying hybrid sampling model:
DepartureTime(tmin = 8, tmax = 10, res = 30) %>% 
  
  #save output in OD_analysis subfolder as My_Departure_Times.dbf
  write.dbf("OD_analysis/My_Departure_Times.dbf")
  
```

## Background

This script was inspired by [Owen & Murphy (2018)](https://trid.trb.org/view/1497217) study and was developed when working on [Stępniak et al. (2019)](https://doi.org/10.1016/j.jtrangeo.2019.01.007) paper:

Stępniak, M., Pritchard, J.P., Geurs K.T., Goliszek S., 2019, *The impact of temporal resolution on public transport accessibility measurement: review and case study in Poland*, Journal of Transport Geography.


### Funding statement

This document is created within the **MSCA CAlCULUS** project.  

*This project has received funding from the European Union's Horizon 2020 research and innovation Programme under the Marie Sklodowska-Curie Grant Agreement no. 749761.*  
*The views and opinions expressed herein do not necessarily reflect those of the European Commission.*

